IIIDYT at IEST 2018: Implicit Emotion Classification With Deep Contextualized Word Representations

### 2018-09-01 paper/paper.tex
[paper] Add brief error analysis, expand conclusion, add pca figure

In this paper we describe our system designed for the WASSA 2018 Implicit
Emotion Shared Task (IEST), which obtained 2$^{\text{nd}}$ place out of 30 teams
with a test macro F1 score of $0.710$. The system is composed of a single
pre-trained ELMo layer for encoding words, a Bidirectional Long-Short Memory
Network BiLSTM for enriching word representations with context, a max-pooling
operation for creating sentence representations from them, and a
Dense Layer for projecting the sentence representations into label space. Our
official submission was obtained by ensembling 6 of these models initialized
with different random seeds.

@@

\textbf{Optimization}: We optimized the model with Adam
\cite{DBLP:journals/corr/KingmaB14}, using default hyperparameters
($\beta_1=0.9$, $\beta_2=0.999$, $\epsilon=10^{-8}$), following a slanted
triangular learning rate schedule \cite{howard2018universal}, also with default
hyperparameters ($cut\_frac=0.1$, $ratio=32$), and a maximum learning rate
$\eta_{max}=0.001$, over $T=23,970$ iterations\footnote{This number is obtained
    by multiplying the number of epochs ($10$) times the total number of
    batches, which for the training dataset corresponds to $2396$ batches of
$64$ elements, and $1$ batch of $39$ elements, hence $2397\times10=23,970$.}.

@@

Once we found the best-performing configuration we trained 10 models using
different random seeds, and tried averaging the output probabilities of all
their possible $\sum_{k=1}^{10}{\binom{10}{k}}=1023$ combinations. As Figure
\ref{fig:best_ensembles} shows, we empirically found that a specific combination
of $6$ models yielded the best results ($70.52\%$), providing evidence for the
fact that using a number of independent classifiers equal to the number of class
labels provides the best results when doing average ensembling
\cite{bonab2016theoretical}.

@@

Figure~\ref{fig:confusion_matrix} shows the confusion matrix of the model
evaluated on the test set, and Table~\ref{table:classification_report} the
corresponding classification report.

@@

@@

As Figure \ref{fig:data_amt_vs_acc} shows, increasing the amount of data in
which our model was trained consistently increased validation accuracy and
validation macro F1 score. The trend shows that having more training data would
likely further increase the model's performance.

@@

>>>>

In this paper we describe our system designed for the WASSA 2018 Implicit
Emotion Shared Task (IEST), which obtained 2$^{\text{nd}}$ place out of 30 teams
with a test macro F1 score of $0.710$. The system is composed of a single
pre-trained ELMo layer for encoding words, a Bidirectional Long-Short Memory
Network BiLSTM for enriching word representations with context, a max-pooling
operation for creating sentence representations from them, and a Dense Layer for
projecting the sentence representations into label space. Our official
submission was obtained by ensembling 6 of these models initialized with
different random seeds. The code for replicating this paper is available at
\url{https://github.com/jabalazs/implicit_emotion}.

@@

\textbf{Optimization}: We optimized the model with Adam
\cite{DBLP:journals/corr/KingmaB14}, using default hyperparameters
($\beta_1=0.9$, $\beta_2=0.999$, $\epsilon=10^{-8}$), following a slanted
triangular learning rate schedule \cite{howard2018universal}, also with default
hyperparameters ($cut\_frac=0.1$, $ratio=32$), and a maximum learning rate
$\eta_{max}=0.001$, over $T=23,970$ iterations\footnote{This number is obtained
    by multiplying the number of epochs ($10$), times the total number of
    batches, which for the training dataset corresponds to $2396$ batches of
$64$ elements, and $1$ batch of $39$ elements, hence $2397\times10=23,970$.}.

@@

Once we found the best-performing configuration we trained 10 models using
different random seeds, and tried averaging the output class probabilities of
all their possible $\sum_{k=1}^{9}{\binom{9}{k}}=511$ combinations. As Figure
\ref{fig:best_ensembles} shows, we empirically found that a specific combination
of $6$ models yielded the best results ($70.52\%$), providing evidence for the
fact that using a number of independent classifiers equal to the number of class
labels provides the best results when doing average ensembling
\cite{bonab2016theoretical}.

@@

Figure~\ref{fig:confusion_matrix} shows the confusion matrix of a single model
evaluated on the test set, and Table~\ref{table:classification_report} the
corresponding classification report. In general, we confirm what
\newcite{klinger2018iest} report: \texttt{anger} was the most difficult class to
predict, followed by \texttt{surprise}, whereas \texttt{joy}, \texttt{fear}, and
\texttt{disgust} are the better performing ones.

To observe whether any particular pattern arose from the sentence
representations encoded by our model, we projected them into 3d space through
Principal Component Analysis (PCA), and were surprised to find that 2 clearly
defined clusters emerged (see Figure \ref{fig:pca}), one containing the majority
of datapoints, and another containing \texttt{joy} tweets exclusively. Upon
further exploration we also found that the smaller cluster was composed only by
tweets containing the pattern \texttt{un~\_\_TRIGGERWORD\_\_}, and further,
that all of them were correctly classified.

It is also worth mentioning that there are 5827 tweets in the training set with
this pattern. Of these, 5822 (99.9\%) correspond to the label \texttt{joy}. We
observe a similar trend on the test set; 1115 of the 1116 tweets having the
\texttt{un~\_\_TRIGGERWORD\_\_} pattern correspond to \texttt{joy} tweets. We
hypothesize this is the reason why the model learned this pattern as a strong
discriminating feature.

Finally, the only tweet in the test set that contained this pattern and did not
belong to the \texttt{joy} class, originally had \textit{unsurprised} as its
triggerword\footnote{We manually searched for the original tweet.}, and
unsurprisingly, was misclassified.

@@

As Figure \ref{fig:data_amt_vs_acc} shows, increasing the amount of data with
which our model was trained consistently increased validation accuracy and
validation macro F1 score. The trend suggests that the proposed model is
expressive enough to learn from more data, and is not overfitting the training
set.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=\columnwidth]{images/pca.pdf}

    \caption{3d Projection of the Test Sentence Representations.}
    \label{fig:pca}
\end{figure}

@@

We also showed that, what was probably an annotation artifact, the
\texttt{un~\_\_TRIGGERWORD\_\_} pattern, resulted in increased performance for
the \texttt{joy} label. This pattern was probably originated by a heuristic
naïvely replacing the ocurrence of \textit{happy} by the trigger-word indicator.
We think the dataset could be improved by replacing the word \textit{unhappy},
in the original examples, by \texttt{\_\_TRIGGERWORD\_\_} instead of
\texttt{un~\_\_TRIGGERWORD\_\_}, and labeling it as \texttt{sad}, or
\texttt{angry}, instead of \texttt{joy}.

% To prevent future models from mistakenly learning this
% feature as a strong indicator for such label, it would be wise to add more
% examples from other classes cotaining the same pattern, or removing them
% entirely and rebalancing the dataset.
